---
title: "Fix bird list prioritization"
output: pdf_document
date: "2023-09-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Add the avian conservation assessment number for conservation 
Here I am going to import the data from the PIF, about the Avian Conservation Assessment Database Scores. Download in: https://pif.birdconservancy.org/avian-conservation-assessment-database-scores/. This file is here named as ACAD

To understand how population declines work for PIF: https://pif.birdconservancy.org/acad.handbook.pdf
Here I am going to use IUCN.Red.List.2018 and population trend "PT.c" for the PIF global. 
For the PIF regional I will use "PT.r". 

ListPIF.csv is going to contain the common name of the species for which you are going to extract the conservation status information

#Extract the IUCN

```{r}
ACAD<-read.csv("C:/Users/Angie/Downloads/ACAD.csv", header = T)
file1<-read.csv("C:/Users/Angie/Downloads/ListPIF.csv", header = T)

for (i in 1: dim(file1)[1]){
  for (j in 1: dim(ACAD)[1]){
    if (file1$CommonName[i]== ACAD$Common.Name[j]){
      file1$IUCN[i]<- ACAD$IUCN.Red.List.2018[j]
      break
    }
  }
}

write.csv(file1,"C:/Users/Angie/Downloads/ListPIF1.csv")
```

#Extract the PT.c
```{r}
ACAD<-read.csv("C:/Users/Angie/Downloads/ACAD.csv", header = T)
file1<-read.csv("C:/Users/Angie/Downloads/ListPIF.csv", header = T)

for (i in 1: dim(file1)[1]){
  for (j in 1: dim(ACAD)[1]){
    if (file1$CommonName[i]== ACAD$Common.Name[j]){
      file1$PT.c[i]<- ACAD$PT.c[j]
      break
    }
  }
}

write.csv(file1,"C:/Users/Angie/Downloads/ListPIF2.csv")
```

Here file 2 is going to contain the PIF regional. To avoid confusion I did not name it ACAD
```{r, echo=FALSE}
file1<-read.csv("C:/Users/Angie/Downloads/ListPIF.csv", header = T)
file2<-read.csv("C:/Users/Angie/Downloads/ListPIFR.csv", header = T)


for (j in 1: dim(ACAD)[1]){
    x<-which(file2$Common.Name==file1$CommonName[j])
      if (length(x)>0){
        for (i in 1:length(x)){
          if (file2$Region[x[i]]=="BCR08"){
            file1$BCR08[j]<-file2$PT.r[x[i]]
          }
          if (file2$Region[x[i]]=="BCR12"){
            file1$BCR12[j]<-file2$PT.r[x[i]]
          }
          if (file2$Region[x[i]]=="BCR14"){
            file1$BCR14[j]<-file2$PT.r[x[i]]
          }
          
        }
   }
     
}



write.csv(file1,"C:/Users/Angie/Downloads/ListPIF3.csv")
```


