---
title: "How to Prioritization"
output:
  html_document:
    df_print: paged
date: "2023-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown is created to document step by step the EHJV prioritization. 


#Step 0. Create a shapefile of the study area (EHJV) using the a shapefile from: https://geoscan.nrcan.gc.ca/starweb/geoscan/servlet.starweb?path=geoscan/downloade.web&search1=R=300911

```{r}
ecosh1<-terra::vect("./BCR_Terrestrial/BCR_Terrestrial_master.shp") #import shape file


ecosh3<-subset(ecosh1, ecosh1$COUNTRY %in% "CANADA")
ecosh3<-subset(ecosh3, ecosh3$PROVINCE_S%in% c("ONTARIO","QUEBEC","NEWFOUNDLAND", "NOVA SCOTIA", "PRINCE EDWARD ISLAND", "NEW BRUNSWICK"))
ecosh3<-subset(ecosh3, ecosh3$BCR%in% c(8,12,14))

plot (ecosh3)
```

```{r}

#save shapefile 

#writeVector(ecosh3, "C:/Angela/BAM_project/Zonation/EHJV/Inputs/EHJVsh.shp", overwrite=TRUE)

```


#Step 1.Clip the national models into the the desire study area 

I created the function [ClipMaps] (https://github.com/angelXmonster/Zonation_EHJV_Prioritization/blob/main/Step1/ClipMaps.R), which clip any raster based on a shapefile containing the study area. 

This function can clip rasters individually or it can be feed with the whole folder where are stored the rasters to be clipped.
(Read the description of the function for more information)

```{r}
#Runs each map
clipMaps ("C:/Angela/BAM_project/Zonation/EHJV/Inputs/6Species/birdRasterFiles/pred-BANS-CAN-Mean.tif", "C:/Angela/BAM_project/Zonation/EHJV/Inputs/EHJVshp/EHJVsh.shp", "C:/Angela/BAM_project/Zonation/EHJV/Inputs/6Species") 

clipMaps("C:/Angela/BAM_project/Zonation/EHJV/Inputs/AllSpecies/National_Models_26_08_2022", "C:/Angela/BAM_project/Zonation/EHJV/Inputs/EHJVshp_plusManitoba/EHJVsh.shp", "C:/Angela/BAM_project/Zonation/EHJV/Inputs/AllSpecies/birdRasterFilesEHJV") 
```

#Step 2. Create the conservation weights

In the Zonation Analyses the species were weighted by their conservation status: 
	5 = Schedule 1 listed species
	4 = Large decreases in population (PIF Pt.c* of 5)
	3 = Moderate decreases in populations (PIF Pt.c of 4)
	2= Priority species by BCR 
	1 = all other species
	
* PIF Pt.c score = North American population trend scores indicating vulnerability due to the direction and magnitude of recent changes in population size. 
* Pt.c score of 5 = Significant large decrease (< -50%); 
  Pt.c score of 4 = Moderate to large decrease ( -15% to -50%)
  See more details in: https://pif.birdconservancy.org/acad.handbook.pdf

Note: For EHJV project I create another RMD file where I extracted the conservation weights coming from PIF. [FixBirdListPrioritization] (https://github.com/angelXmonster/Zonation_EHJV_Prioritization/blob/main/Step1/FixBirdListPrioritization.Rmd).

#Step 3. Add AUC weight 

Before adding the AUC weight please note that all national models with AUC lower than 0.7 are discarded. 
Also any national model that does not have AUC. 
Note: To check the national models AUC please go to [BAM/Methods] (https://borealbirds.github.io/methods/) and download the result.xlsx

First we rescale the conservation weight as well as the AUC weight. So, both go from 0 to 1.

To rescale the AUC scores: AUC' = (AUC - 0.5) * 2. 

To rescale the Conservation scores: CS' = CS/5

Then average: (a*AUC' + b*CS')/(a+b). Where a and b are the "weights" one whats to gives to the AUC and CS.
In this case both are equally weight so it is going to be 1. 

```{r}

```

#Step 4. Write the zonation features file

In zonation, each species raster is called feature.
To run zonation one need to create a text file with the directory of each feature and the weight. 

I will create a list containing the names of the raster files, all of which are going to be in a single folder.
Such rasters need to be in the form: pred-(ID)-CAN-MeanClipped.tif
I will need a csv with the birds ID and their respective weights 

In this case I am only using landbirds

##make a list from the folder 

Here I are creating a list of the rasters to be used in zonation

```{r}
r<-"C:/Angela/BAM_project/Zonation/EHJV/Inputs/AllSpecies/birdRasterFilesEHJV/Land"
listb<-list.files(r)
```

create a list with only the names. This is going to be used to match the birds IDson the csv file 
```{r}
nam<-listb
for (i in 1:length(listb)){
nam[i]<- gsub('-CAN-MeanClipped.tif','',listb[i])
nam[i]<- gsub('pred-','',nam[i])
}
```

##Create the text file

dtxt is going to be a list with the IDs and Weights of the birds extracted from the csv file

```{r}
dtxt<-read.csv("C:/Angela/BAM_project/Zonation/EHJV/Inputs/SpeciesList1.csv", header = T)

str(dtxt) #chech the data looks the way you want 

tx<-dtxt %>% select(ConservationWeight, ID)


for (j in 1:dim(tx)[1]){
   x<-which(nam==tx$ID[j])
   if (length(x)>0){
   preloc<- paste (r, "/", sep = "")
    tx$Dir[j]<-paste (preloc, listb[x[1]], sep = "")
   }
   else{ tx$Dir[j]<-NA}
}

tx<-tx %>% select(ConservationWeight, Dir)
tx<-tx [which (!is.na(tx$Dir)),]

colnames(tx) <- c("weight","filename")

#write.table(tx, "C:/Angela/BAM_project/Zonation/EHJV/Inputs/features_SpeciesList1.txt", quote = T, row.names = FALSE)
```

#Step 5. Run Prioritization

#Step 6. Plot prioritization 

##Compare with protected areas 
#Step 7. Include more variables
## Include footprint
## Include intacness
